---
title: "R Notebook"
output: html_notebook
---
Whats in V4:
GAM models using Humidity, Temperature, PM2.5, NO2 and NPIs

Climate variables (temperature, humidity, PM2.5) lagged so that daily cases are predicted using climate variables from n=climate_lag days prior

NPI lagged so that daily cases are predicted using NPI data from n=0,...,num_iter days prior

the climate splines should be more or less the same since we're keeping climate_lag constant
i think what we're looking for is an appropriate lag for npi which should change a lot over the num_iter GAMs

If you run this you'll see that infections decrease with temperature and increase with PM2.5
Humidity is kind of a wildcard. 
This is overall true for all the cities (florence is a little weird) and for all the values of climate lag I tried (3,4,5,6)


We start by loading in our data tables

```{r}
require(mgcv)
bologna <- read.csv("bologna_full.csv")
bologna <- na.omit(bologna)

modena <- read.csv("modena_full.csv")
modena <- na.omit(modena)

florence <- read.csv("florence_full.csv")
florence <- na.omit(florence)

rome <- read.csv("rome_full.csv")
rome <- na.omit(rome)

num_lag <- 21
climate_lag <- 4
```


Bologna:
```{r}
for(i in num_lag:num_lag) {
num_lag <- i
bologna <- read.csv("bologna_full.csv") #Read in file
#Push temperature&humidity data back climate_lag days by adding empty data in front
#rep(NA, n) creates a vector of length n filled with NA
bologna$Temperature.1 <- c(rep(NA, climate_lag),  bologna$Temperature[(1):(nrow(bologna)-climate_lag)]) 
bologna$Humidity.1 <- c(rep(NA, climate_lag), bologna$Humidity[(1):(nrow(bologna)-climate_lag)])
bologna$pm25.1 <- c(rep(NA, climate_lag), bologna$pm25[(1):(nrow(bologna)-climate_lag)])
bologna$no2.1 <- c(rep(NA, climate_lag), bologna$no2[(1):(nrow(bologna)-climate_lag)])


#push npi back num_lag days
bologna$npi.1 <- c(rep(NA, num_lag), bologna$npi[(1):(nrow(bologna)-num_lag)])
#Omit the new NAs
bologna <- na.omit(bologna)

infections.gam <- gam(Daily ~ s(Temperature.1) + s(Humidity.1) + s(pm25.1) + s(no2.1)+ s(npi.1, k=4), family=poisson(link="log"), data=bologna)
plot(infections.gam,scale=0,se=2, shade=TRUE,pages=1)
title(main = i)
}
```

Modena:
```{r}
for(i in num_lag:num_lag) {
num_lag <- i
modena <- read.csv("modena_full.csv") #Read in file
#Push temperature&humidity data back climate_lag days by adding empty data in front
#rep(NA, n) creates a vector of length n filled with NA
modena$Temperature.1 <- c(rep(NA, climate_lag),  modena$Temperature[(1):(nrow(modena)-climate_lag)]) 
modena$Humidity.1 <- c(rep(NA, climate_lag), modena$Humidity[(1):(nrow(modena)-climate_lag)])
modena$pm25.1 <- c(rep(NA, climate_lag), modena$pm25[(1):(nrow(modena)-climate_lag)])
modena$no2.1 <- c(rep(NA, climate_lag), modena$no2[(1):(nrow(modena)-climate_lag)])

#push npi back num_lag days
modena$npi.1 <- c(rep(NA, num_lag), modena$npi[(1):(nrow(modena)-num_lag)])
#Omit the new NAs
modena <- na.omit(modena)

infections.gam <- gam(Daily ~ s(Temperature.1) + s(Humidity.1) + s(pm25.1) + s(no2.1) + s(npi.1, k=4), family=poisson(link="log"), data=modena)
plot(infections.gam,scale=0,se=2, shade=TRUE,pages=1)
title(main = i)
}

```




Rome:
```{r}
for(i in num_lag:num_lag) {
num_lag <- i
rome <- read.csv("rome_full.csv") #Read in file
#Push temperature&humidity data back climate_lag days by adding empty data in front
#rep(NA, n) creates a vector of length n filled with NA
rome$Temperature.1 <- c(rep(NA, climate_lag),  rome$Temperature[(1):(nrow(rome)-climate_lag)]) 
rome$Humidity.1 <- c(rep(NA, climate_lag), rome$Humidity[(1):(nrow(rome)-climate_lag)])
rome$pm25.1 <- c(rep(NA, climate_lag), rome$pm25[(1):(nrow(rome)-climate_lag)])
rome$no2.1 <- c(rep(NA, climate_lag), rome$no2[(1):(nrow(rome)-climate_lag)])

#push npi back num_lag days
rome$npi.1 <- c(rep(NA, num_lag), rome$npi[(1):(nrow(rome)-num_lag)])
#Omit the new NAs
rome <- na.omit(rome)

infections.gam <- gam(Daily ~ s(Temperature.1) + s(Humidity.1) + s(pm25.1)+s(npi.1, k=4) + s(no2.1), family=poisson(link="log"), data=rome)
plot(infections.gam,scale=0,se=2, shade=TRUE,pages=1)
title(main = i)
}
```
