---
title: "GAM Example Milan"
output:
  pdf_document: default
  word_document: default
---

```{r}
library(mgcv)
library(nlme)
```

Here, we describe fitting a GAM to COVID-19 and environmental data from Milan, Italy.
First, we read in the .csv file while omitting null values.

```{r}
milan_full <- read.csv("milan_full.csv")
milan_full <- na.omit(milan_full)
```

Then, we create the GAM. RR relative rate of infection (the actual rate of infection divided by the average rate of infection), and we have created a three-day lag behind reported data. We want to see how humidity, temperature, PM2.5, PM10, NO2, and O3 influence the relative rate. We use a gaussian distribution for the GAM with a log link function.

```{r}
infections.gam <- gam(LAGRR ~ s(Humidity) + s(Temperature) + s(pm25) + s(pm10) + s(no) + s(o3), family=gaussian(link="log"), data=milan_full)
```

We plot each of the individual splines, which show how each variable influences relative rate if all other variables are held constant.

```{r}
plot(infections.gam, ylim=c(-1,1),scale=0,se=2, residuals=TRUE, shade=TRUE,pages=1)
```

Finally, we print a summary of the GAM's results.

```{r}
summary(infections.gam)
```

