---
title: "GAM Example Milan"
output:
  pdf_document: default
  word_document: default
---

```{r}
library(mgcv)
library(nlme)
```

Here, we describe fitting a GAM to COVID-19 and environmental data from Milan, Italy.
First, we read in the .csv file while omitting null values.

```{r}
milan_full <- read.csv("milan_full.csv")
milan_full <- na.omit(milan_full)
```

Then, we create the GAM. RR relative rate of infection (the actual rate of infection divided by the average rate of infection), and we have created a three-day lag behind reported data. We want to see how humidity, temperature, PM2.5, PM10, NO2, and O3 influence the relative rate. We use a gaussian distribution for the GAM with a log link function.

```{r}
infections.gam <- gam(LAGRR ~ s(Humidity) + s(Temperature) + s(pm25) + s(pm10) + s(no) + s(o3), family=gaussian(link="log"), data=milan_full)
```

We plot each of the individual splines, which show how each variable influences relative rate if all other variables are held constant.

```{r}
plot(infections.gam, ylim=c(-1,1),scale=0,se=2, residuals=TRUE, shade=TRUE,pages=1)
```

Finally, we print a summary of the GAM's results.

```{r}
summary(infections.gam)
```
Plotting the residuals: 
```{r}
par(mfrow = c(2,2))
gam.check(infections.gam)
```
Plot of true vs predicted values (same as 4th plot above)
```{r}
pred_infectionsgam = predict(infections.gam, type = "response")
plot(pred_infectionsgam, milan_full$LAGRR, xlab = "Predicted LAGRR", ylab = "True LAGRR")
abline(0,1)
```
Here, I included a smooth for multiple predictors (Temperuate and Humidity to start as they seem like thye could be related). The first method uses a tensor product spline fit, the second method uses a thin plate regression spline fit. I don't have any intuition on how to interpret the plots when smoothing two predictors at a time yet but will look into it more..

```{r}
infections.gamTE <- gam(LAGRR ~ te(Humidity,Temperature) + s(pm25) + s(pm10) + s(no) + s(o3), family=gaussian(link="log"), data=milan_full)
summary(infections.gamTE)
plot(infections.gamTE)
```

Thin plate splines:
```{r}
infections.gamtps <- gam(LAGRR ~ s(Humidity,Temperature) + s(pm25) + s(pm10) + s(no) + s(o3), family=gaussian(link="log"), data=milan_full)
summary(infections.gamtps)
plot(infections.gamtps)
```
Note that the GCV for the method using tensor product splines is lower than for TPS and the adjusted R^2 is higer for tensor splines as well. 


We can use ANOVA to do a quick side-by-side comparison of the three methods:
```{r}
anova.gam(infections.gamTE, infections.gamtps, infections.gam)
```
Comparison of AIC scores:
```{r}
AIC(infections.gamtps, infections.gamTE, infections.gam)
```

